---
title: "WHuang PBMC project"
author: "Jiangyan Yu (jiangyan.yu.de@gmail.com)"
date: "`r Sys.Date()`"
output:
  html_document: 
    code_folding: hide
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
in current analysis, no batch correction was performed.

# general steps

```{r global_options}
knitr::opts_chunk$set(warning=FALSE, messgae=FALSE, fig.path='Figs/', results = "hide")
## fig.width=4, fig.height=4
```

## load library

```{r include=FALSE}
rm(list=ls())
gc()
#CRAN packages
list.of.packages <- c("readr",
                      "cowplot",
                      "useful",
                      "stringr",
                      "umap",
                      "ggplot2",
                      "reshape2",
                      "dplyr",
                      "tidyr",
                      "Matrix.utils",
                      "VGAM",
                      "plotly",
                      "future",
                      "data.table"
)

# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
# if(length(new.packages)>0) 
#   install.packages(new.packages)

#BioconductoR packages
list.of.bioc.packages <- c("tximport",
                           "DESeq2",
                           "Seurat",
                           "slingshot",
                           "flowCore",
                           "biomaRt",
                           "clusterProfiler",
                           "org.Hs.eg.db",
                           "org.Mm.eg.db",
                           "GSEABase",
                           "DOSE",
                           "BiocGenerics",
                           "DelayedArray",
                           "DelayedMatrixStats",
                           "limma",
                           "S4Vectors",
                           "SingleCellExperiment",
                           "SummarizedExperiment",
                           "batchelor",
                           "ComplexHeatmap")
                           # "annotables")
# new.packages.bioc <- list.of.bioc.packages[!(list.of.bioc.packages %in% installed.packages()[,"Package"])]
# # 
# if(length(new.packages.bioc)>0)if (!requireNamespace("BiocManager")) install.packages("BiocManager")
# BiocManager::install(new.packages.bioc, update = FALSE)

lapply(c(list.of.packages,list.of.bioc.packages), require, character.only = TRUE)

rm(list.of.bioc.packages,list.of.packages,new.packages,new.packages.bioc)
```

## working directory

```{r}
## Mac directory
working.dir = "/home/jyu/rstudio/"
working.dir = "/Users/jiangyanyu/sciebo/Projects2023/Dafsari_PUF60_ownerJYu/"
global_ref_dir = paste0(working.dir,"/analysis_jyu/Scripts/")
gsea_pathway_dir = paste0(working.dir,"/analysis_jyu/Scripts/")
# source(paste0(global_ref_dir,"general_functions.R"))
# 
allcolour=c("#DC143C","#0000FF","#20B2AA","#FFA500","#9370DB","#98FB98","#F08080","#1E90FF","#7CFC00","#FFFF00",
              "#808000","#FF00FF","#FA8072","#7B68EE","#9400D3","#800080","#A0522D","#D2B48C","#D2691E","#87CEEB","#40E0D0","#5F9EA0",
              "#FF1493","#0000CD","#008B8B","#FFE4B5","#8A2BE2","#228B22","#E9967A","#4682B4","#32CD32","#F0E68C","#FFFFE0","#EE82EE",
              "#FF6347","#6A5ACD","#9932CC","#8B008B","#8B4513","#DEB887")
```


# load save object

```{r}
vdj_combined = readRDS(file = paste0(working.dir,"/analysis_jyu/vdj_8combined_1kfeature_noBatchRmv1.rds"))
```

## add clinical info

```{r}
sample_info = read.csv(file=paste0(working.dir,"/analysis_jyu/sample_info.txt"))
tmp_data = vdj_combined@meta.data
tmp_data$order = c(1:nrow(tmp_data))
tmp_data = merge(tmp_data[,c("sample","order")],sample_info,by.x="sample",by.y="sample")
tmp_data = tmp_data[order(tmp_data$order),]
rownames(tmp_data) = rownames(vdj_combined@meta.data)

vdj_combined = AddMetaData(vdj_combined,metadata = tmp_data[,"variant"],col.name = "variant")
vdj_combined = AddMetaData(vdj_combined,metadata = tmp_data[,"age"],col.name = "age")
vdj_combined = AddMetaData(vdj_combined,metadata = tmp_data[,"relationship"],col.name = "relationship")

rm(sample_info, tmp_data)
```
## rename cells
ClassicalMonocyte, Megakaryocyte

0:restingCD4+ (IL7R,AQP3) source: https://www.nature.com/articles/s41467-019-12464-3
1,5:NaiveCD4+(SELL,KLF2,TCF7): https://www.nature.com/articles/s41467-019-12464-3
11:Treg/Tfh (FXP3,CTLA4):https://www.nature.com/articles/s41467-019-12464-3
3:cytotoxicCD8+Tcell(FCGR3A,KLRF1): https://www.cell.com/action/showPdf?pii=S0092-8674%2818%2931568-X
4:naiveCD8+(CCR7,SELL,TCF7):https://www.nature.com/articles/s41467-019-12464-3
10,15: dysfunctionCD8(LAG3):https://www.cell.com/action/showPdf?pii=S0092-8674%2818%2931568-X
14:gamma-delta (TRDV2, TRGV9): https://www.nature.com/articles/s41590-020-0762-x

### cluster_celltype

```{r}
Idents(vdj_combined) = "seurat_clusters"

vdj_combined = RenameIdents(vdj_combined,
                            "0" = "0:RestingCD4+Tcell", 
                            "1" = "1:NaiveCD4+Tcell",
                            "2" = "2:CD8+Tcell",
                            "3" = "3:CytotoxicCD8+Tcell",
                            "4" = "4:NaiveCD8+Tcell",
                            "5" = "5:NaiveCD4+Tcell",
                            "6" = "6:NKcell",
                            "7" = "7:Bcell",
                            "8" = "8:CD8+Tcell",
                            "9" = "9:Monocyte",
                            "10" = "10:DysfunctionCD8+Tcell",
                            "11" = "11:Tregcell",
                            "12" = "12:Bcell",
                            "13" = "13:Doublets",
                            "14" = "14:GammaDeltaTcell",
                            "15" = "15:DysfunctionCD8+Tcell",
                            "16" = "16:Megakaryocyte",
                            "17" = "17:Bcell",
                            "18" = "18:NonClassicalMonocyte",
                            "19" = "19:NKcell",
                            "20" = "20:NaiveCD4+Tcell",
                            "21" = "21:pDC",
                            "22" = "22:Plasmablasts"
                            )

vdj_combined$cluster_celltype = vdj_combined@active.ident
```

### cell type

```{r}
Idents(vdj_combined) = "seurat_clusters"

vdj_combined = RenameIdents(vdj_combined,
                            "0" = "CD4+Tcell",
                            "1" = "CD4+Tcell",
                            "2" = "CD8+Tcell",
                            "3" = "CD8+Tcell",
                            "4" = "CD8+Tcell",
                            "5" = "CD4+Tcell",
                            "6" = "NKcell",
                            "7" = "Bcell",
                            "8" = "CD8+Tcell",
                            "9" = "Monocyte",
                            "10" = "CD8+Tcell",
                            "11" = "CD4+Tcell",
                            "12" = "Bcell",
                            "13" = "CD8+Tcell",
                            "14" = "CD8+Tcell",
                            "15" = "CD8+Tcell",
                            "16" = "Megakaryocyte",
                            "17" = "Bcell",
                            "18" = "NonClassicalMonocyte",
                            "19" = "NKcell",
                            "20" = "CD4+Tcell",
                            "21" = "pDC",
                            "22" = "Plasmablasts")

vdj_combined$celltype = vdj_combined@active.ident

Idents(vdj_combined) = "seurat_clusters"

vdj_combined = RenameIdents(vdj_combined,
                            "0" = "Tcell",
                            "1" = "Tcell",
                            "2" = "Tcell",
                            "3" = "Tcell",
                            "4" = "Tcell",
                            "5" = "Tcell",
                            "6" = "NKcell",
                            "7" = "Bcell",
                            "8" = "Tcell",
                            "9" = "Monocyte",
                            "10" = "Tcell",
                            "11" = "Tcell",
                            "12" = "Bcell",
                            "13" = "NKcell",
                            "14" = "Tcell",
                            "15" = "Tcell",
                            "16" = "Megakaryocyte",
                            "17" = "Bcell",
                            "18" = "Monocyte",
                            "19" = "NKcell",
                            "20" = "Tcell",
                            "21" = "pDC",
                            "22" = "Plasmablasts")

vdj_combined$major = vdj_combined@active.ident
```

# select OAS2 patients only

```{r}
oas2 = subset(vdj_combined,subset=variant=="OAS2")
rm(vdj_combined)
gc()
```

## label texts

```{r}
cluster_anno = c("0" = "0:CD4+Tcell",
                            "1" = "1:CD4+Tcell",
                            "2" = "2:CD8+Tcell",
                            "3" = "3:CD8+Tcell",
                            "4" = "4:CD8+Tcell",
                            "5" = "5:CD4+Tcell",
                            "6" = "6:NKcell",
                            "7" = "7:Bcell",
                            "8" = "8:CD8+Tcell",
                            "9" = "9:Monocyte",
                            "10" = "10:CD8+Tcell",
                            "11" = "11:CD4+Tcell",
                            "12" = "12:Bcell",
                            "13" = "13:CD8+Tcell",
                            "14" = "14:CD8+Tcell",
                            "15" = "15:CD8+Tcell",
                            "16" = "16:Megakaryocyte",
                            "17" = "17:Bcell",
                            "18" = "18:NonClassicalMonocyte",
                            "19" = "19:NKcell",
                            "20" = "20:Plasmablasts",
                            "21" = "21:pDC",
                            "22" = "22:Plasmablasts")
```


# fig1:overall umap 

```{r}
fig1 = DimPlot(oas2,group.by = "seurat_clusters",dims = c(1,3),label = TRUE)+
  labs(title = c(""))+
  geom_text(x=10,y=-5,label = "n = 10,671 cells",check_overlap = TRUE)+
  # theme(legend.position = "NULL")+
  scale_color_discrete(name = "", labels = cluster_anno)+
  theme(axis.text = element_text(size = 10))

pdf(file = paste0(working.dir,"/AOS2_paper/fig1_overall_umap.pdf"),height = 6,width = 10)
fig1
dev.off()
```


# figS1-3:quality vlnplot

## s1: feature_RNA

```{r}
figs1 = VlnPlot(oas2,features = c("nFeature_RNA"),group.by = "cluster_celltype",pt.size = 0)+
  coord_flip()+
  labs(title = c("Number of unqiue genes"))+
  xlab("")+
  # geom_text(x=10,y=-5,label = "n = 10,671 cells")+
  theme(legend.position = "NULL",
        plot.title = element_text(size = 10,face = "plain"),
        axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5),
        axis.text = element_text(size = 10))
  # scale_color_discrete(name = "", labels = cluster_anno)

pdf(file = paste0(working.dir,"/AOS2_paper/figs1_nFeature_RNA.pdf"),height = 6,width = 5)
figs1
dev.off()
```

## s2:mt_percent

```{r}
figs2 = VlnPlot(oas2,features = c("percent.mt"),group.by = "cluster_celltype",pt.size = 0)+
  coord_flip()+
  labs(title = c("Frequency of mitochondria genes (%)"))+
  xlab("")+
  # geom_text(x=10,y=-5,label = "n = 10,671 cells")+
  theme(legend.position = "NULL",
        plot.title = element_text(size = 10,face = "plain"),
        axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5),
        axis.text = element_text(size = 10))

pdf(file = paste0(working.dir,"/AOS2_paper/figs2_percent_mt.pdf"),height = 6,width = 5)
figs2
dev.off()
```

## s3: umap split by sample

```{r}
figs3_left = DimPlot(subset(oas2,subset=age==17),dims = c(1,3),label = TRUE,group.by = "seurat_clusters")+
  labs(title = c("HD0011,17 yrs (7,787 cells)"))+
  theme(legend.position = "NULL",
        plot.title = element_text(size = 10,face = "plain"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10))


figs3_right = DimPlot(subset(oas2,subset=age==23),dims = c(1,3),label = TRUE,group.by = "seurat_clusters")+
  labs(title = c("HD0025,23 yrs (2,884 cells)"))+
  theme(legend.position = "NULL",
        plot.title = element_text(size = 10,face = "plain"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10))


pdf(file = paste0(working.dir,"/AOS2_paper/figs3_umap_pslit_by_sample.pdf"),height = 3,width = 6)
figs3_left + figs3_right
dev.off()

```


# fig2: canonical markers

```{r}
cluster_order = c("0:RestingCD4+Tcell", "1:NaiveCD4+Tcell","5:NaiveCD4+Tcell","20:NaiveCD4+Tcell","2:CD8+Tcell","3:CytotoxicCD8+Tcell","4:NaiveCD8+Tcell","8:CD8+Tcell","10:DysfunctionCD8+Tcell","14:GammaDeltaTcell","11:Tregcell","6:NKcell","19:NKcell","7:Bcell","12:Bcell","17:Bcell","9:Monocyte","18:NonClassicalMonocyte","21:pDC","16:Megakaryocyte","22:Plasmablasts","13:Doublets")

canonical_markers = c("PTPRC","CD3D","CD4","IL7R","AQP3","CCR7","TCF7","SELL","CD8B","CD8A","LAG3","TRDV2","TRGV9","CTLA4","FOXP3","GNLY","NCR1","CD19","MS4A1","CD14","FCGR3A","FLT3","PPBP","PF4","IGLC2","IGHA1")
fig2 = DotPlot(oas2,features = canonical_markers,group.by = "cluster_celltype")+
  scale_y_discrete(limits = cluster_order) +
  theme(axis.text.x = element_text(size = 10,angle = 90, vjust = 0.5, hjust=1),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10))


pdf(file = paste0(working.dir,"/AOS2_paper/fig2_dotplot_canonical_markers.pdf"),height = 6,width = 12)
fig2
dev.off()
```

# figS4: deg heatmap

```{r}
deg = read.csv(file = paste0(working.dir,"/analysis_jyu/vdj_8combined_1kfeature_noBatchRmv_deg.csv"))

deg %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10

plot_cells = oas2@meta.data
plot_cells$cell = rownames(oas2@meta.data)
plot_cells %>% 
  group_by(seurat_clusters) %>% 
  slice_sample(n = 20) -> cells


figs4 = DoHeatmap(oas2[,cells$cell], features = top10$gene,group.by = "seurat_clusters",angle = 90) + NoLegend()

jpeg(file = paste0(working.dir,"/AOS2_paper/figs4_deg_heatmap.jpeg"),height = 1600,width = 600,quality = 300)
figs4
dev.off()

rm(plot_cells,cells,deg,top10)
```

# fig3: cell frequency

## function

```{r}
cluster_freq = function(seurat_object,sample_meta,cell_type,selected_sample){
  seurat_object = seurat_object
  sample_meta = sample_meta
  cell_type = cell_type
  selected_sample = selected_sample

  tmp_data = cbind(sample=seurat_object@meta.data[,sample_meta],cell_type= paste0("",seurat_object@meta.data[,cell_type] )) %>% as.data.frame()
  ns <- table(sample = tmp_data$sample, Type = tmp_data$cell_type)
  fq <- prop.table(ns, 1) * 100
  df <- as.data.frame(fq)
  df = subset(df,sample==selected_sample)
  print(df)
  return(df)
}

freq_plot=function(df){
  ## barplot
  
  # p = ggplot(df,aes(x=sample,y=Freq,fill=cell_type))+
  #       geom_bar(stat = "identity")+
  #   theme_classic()+
  #       theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
  ## donut plot
  data = df
  # data$cell_type = factor(data$cell_type,levels = levels(paste0("",seurat_object@meta.data[,cell_type] )))
  # data = data[order(data$Type), ]
  # data$Freq = round(data$Freq, digits = 1)
  # # Compute the cumulative percentages (top of each rectangle)
  # data$ymax <- cumsum(data$Freq)
  # 
  # # Compute the bottom of each rectangle
  # data$ymin <- c(0, head(data$ymax, n = -1))
  # 
  # # Compute label position
  # data$labelPositionY <- (data$ymax + data$ymin) / 2
  # # data$labelPositionX <- (data$xmax) / 2
  # # Compute a good label
  # data$label <- paste0(data$cell_type, "\n value: ", data$Freq)
  
  # Make the plot
  hsize = 2
  p = ggplot(data, aes(
    x = hsize,
    y = Freq,
    fill = Type,
    label = Type
  )) +
    geom_col(color = "black") +
    # ggrepel::geom_text_repel(position = position_stack(vjust = 0.5),direction = "y", vjust = 1,size = 4.5,min.segment.length = 6)+
    # geom_text(aes(label = cell_type),
    #           position = position_stack(vjust = 0.5)) +
    coord_polar(theta = "y") +
    # scale_fill_brewer(palette = "GnBu") +
    xlim(c(0.2, hsize + 0.5)) +
    theme(
      panel.background = element_rect(fill = "white"),
      panel.grid = element_blank(),
      axis.title = element_blank(),
      axis.ticks = element_blank(),
      axis.text = element_blank()
    )
  
  plot(p)

  return(p)

}
```

## plot

```{r}
fig3_left = cluster_freq(seurat_object = oas2,
                      sample_meta = "sample",
                      cell_type = "major",
                      selected_sample = "HD0011_PBMC") %>% freq_plot()

fig3_right = cluster_freq(seurat_object = oas2,
                      sample_meta = "sample",
                      cell_type = "major",
                      selected_sample = "HD0025_PBMC") %>% freq_plot()

pdf(file = paste0(working.dir,"/AOS2_paper/fig3_cell_frequency.pdf"),height = 3,width = 5)
fig3_left+ labs(title = "HD0011")+
  theme(plot.title = element_text(size = 10,face = "plain",hjust=0.5),
        legend.position = "NULL")+
  fig3_right+ labs(title = "HD0025")+
  theme(plot.title = element_text(size = 10,face = "plain",hjust=0.5))
dev.off()
```

# session infor

```{r}
sessionInfo()
```

